---
title: "Viz_Part1"
author: "Jesus Luevano"
date: "2023-10-01"
output: github_document
---

```{r}
library(tidyverse)
library(ggridges)

#Options to set figure wedith/sets for WHOLE RMD
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6, #keep same ratio based on above width
  out.width = "90%" #take up 90% width of html doc
)
```

Get the data for plotting

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(  #rnoa pulls in specific dataset from weather
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode( #can also consider case_match or case_when
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10, #to make whoel degree Celsius
    tmax = tmax / 10) |>
  select(name, id, everything())
```


Let's make a plkot 
```{r}
ggplot(weather_df, aes(x = tmin, y= tmax)) + #input data, assign axis (but doesn't pull in how to plot)
       geom_point() #decide to give each datapoint a point on plot

weather_df %>% # can pipe in from dyplr
  ggplot(aes(x = tmin, y = tmax)) + geom_point()

weather_df %>% # can pipe in from dyplr
  filter(name == "CentralPark_NY") %>%
  ggplot(aes(x = tmin, y = tmax)) + geom_point()

#save as varibale
ggp_nycweather =
  weather_df %>% 
  filter(name == "CentralPark_NY") %>%
  ggplot(aes(x = tmin, y = tmax)) + geom_point()

```


## Fancy plot tools

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point() +
  geom_smooth() # fits in line through each colored variable as in ggplot command

ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color=name)) +
  geom_smooth() # fits in line through each colored variable, but as color in geom_point only, lines based on info put in on primary ggplot command

ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point(aes(color = name), alpha = 0.3) + # to make them a bit more translucent
  geom_smooth(se = FALSE) #standard error removed
```

## Plot with facets
```{r}
ggplot(weather_df, aes(x= tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.3) + #alpha blending
  geom_smooth() +
  facet_grid(. ~ name) #one row, separated by name, or name on one row, vs facet_grid( name ~.), generally orders variable based on alphabetical order (but can change up using factors)
```

Let's try a different plot. Temps are boring

```{r}
ggplot(weather_df, aes(x= date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = 0.3) +
  geom_smooth() + 
  facet_grid(. ~ name)
```

Try assigning a specific color
```{r}
weather_df %>%
  filter(name == "CentralPark_NY") %>%
  ggplot(aes(x = date, y = tmax)) + #taking data and mapping aestehtics to plot
  geom_point(color = "blue") # modifying details of presentation only

weather_df %>%
  filter(name != "CentralPark_NY") %>%
  #rename(Date = date)
  ggplot(aes(x = date, y = tmax, color = name)) + #taking data and mapping aestehtics to plot
  geom_point(alpha = 0.7, size = 0.5) # modifying details of presentation only
```


```{r}
weather_df %>%
  ggplot(aes(x = tmin, y= tmax)) + 
  geom_hex() #make some heat mapping to data points on plot when lots of data to show
```


Line plots!
```{r}
#line plots, for longitudinal dataset
weather_df %>%
  filter(name == "Molokai_HI") %>%
  ggplot(aes(x = date, y = tmax)) + 
  geom_line(alpha = 0.5) + 
  geom_point(size = 0.5)
```


## Univariate plotting

Histogram
```{r}
ggplot(weather_df, aes(x= tmax, color = name)) +
  geom_histogram()

ggplot(weather_df, aes(x= tmax, fill = name)) + 
  geom_histogram()  #5

ggplot(weather_df, aes(x = tmax, fill = name)) +
  geom_histogram(position = "dodge")
```


Let's use density plot
```{r}
ggplot(weather_df, aes(x = tmax, fill = name)) +
  geom_density(alpha = 0.3, adjust = 0.5) #sort of smoothed over histogram, alpha for transparency, adjust for smoothness (avoid over- or under-smooothing)
```

Using Boxplots!
```{r}
ggplot(weather_df, aes(y = tmax)) + #so vertical
  geom_boxplot()

ggplot(weather_df, aes(y = tmax, x= name)) + #so vertical y first, then x to break up
  geom_boxplot()

```

Violin plots?
```{r}
ggplot(weather_df, aes(y = tmax, x= name)) +
  geom_violin() #orient vertically then flip
```

Ridge plot
```{r}
ggplot(weather_df, aes(x = tmax, y = name)) +
  geom_density_ridges()
```

## Saving and embedding plots

```{r}
ggp_weather =
  weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()

ggp_weather

ggsave("data/Results/ggp_weather.pdf", ggp_weather)
```

```{r, fig.width= 12, fig.height= 12}
ggp_weather

```

```{r, fig.width= 6, fig.asp= 0.6}
ggp_weather
```

# Learning assessment
 Make plots that compare precipitation across locations. Try a histogram, a density plot, a boxplot, a violin plot, and a ridgeplot; use aesthetic mappings to make your figure readable.
 
```{r}
weather_df %>%
  ggplot(aes(x = prcp)) + 
  geom_histogram(aes(color = name))

#density plot 
ggplot(weather_df, aes(x = prcp)) + 
  geom_density(aes(fill = name), alpha = .5) 

#ridge plot
ggplot(weather_df, aes(x = prcp, y = name)) + 
  geom_density_ridges(scale = .85)

#boxplot
ggplot(weather_df, aes(y = prcp, x = name)) + 
  geom_boxplot() 

weather_df |> 
  filter(prcp > 0) |> 
  ggplot(aes(x = prcp, y = name)) + 
  geom_density_ridges(scale = .85)

```
 
